The Etiology of Cybersecurity

### 2021-03-02 reports/paper_0/main.tex
remove last excel ref

Our literature review shows that most of the definitions relate insecurity to
dis-honesty (also called maliciousness or adversarial) of an agent (often
called adversary or attacker). In fact, weaknesses become vulnerability if an
attacker exploits them in an attack.  This, however, just shifts the problem of
defining what cybersecurity is as the problem of defining what a dishonest
agent can or cannot do in a system, where an agent is any virtual or physical
entity of the system or using the system (e.g. a device, a software, or a human
being), and dishonesty is not necessarily related to malicious motivation but
also to incompetence or lack of skills. As in the Dolev-Yao attacker model\footnote{The Dolev-Yao
attacker model can be considered as abstract model of an attacker that can be
used for the analysis of protocol specifications.} \autocite{Dolev1983security}, we may
correlate ``being dishonest'' to ``not following the intended behavior/rules''.
In the case of a generic system, a dishonest agent is, therefore, any agent
that doesn't follow the intended behavior (or functionality or logic) of the
system.  For example, a software can be considered an agent of the system and
whenever it has a bug, it can be exploited causing an incident. This
may lead to the conclusion that the dishonest behaviors of agents cannot be
defined in general (e.g. due to the heterogeneity of agents and systems) and
that a huge repository of dishonest behaviors should be kept as definition, such
as CAPEC \autocite{MITRE2020CAPEC}; or that dishonesty of an agent with respect
to a cybersecurity protocol should be defined as a number of predefined actions as
in
\autocite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}
(to name a few)\footnote{In formal protocol verifiers, in the DY model, the
attacker has a set of hardcoded inference rules that defines how the attacker
deduces new messages based on the observed message exchange.}.  As depicted in
Figure~\ref{fig:causality}, in order to define a theory on cybersecurity we may
say that the presence of vulnerabilities is a necessary condition in order to
cause an attack in the system.  Those vulnerabilities are, in turn, caused by
the presence of weaknsesses in the system.  Weaknesses are errors in the design
or implementation of a system.  Therefore, a theory on cybersecurity should
first predict the errors in a system design.

@@

%and the information or 

@@

% & \cellcolor{abfred}& %dr
%	\cellcolor{abf-rg-1}& %\pp{$\assertionRegion$}{$\factRegion$} & %po
%	\cellcolor{abf-rg-2}  & %pp
%	\cellcolor{abf-rg-3} & %\pp{$\assertionRegion$}{$\factRegion$} & %ppi
%	\cellcolor{abf-rg-4} \\ %eq

@@

\subsection{From Multi-Agent to Cyber-Physical Systems}
\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth]{abftheory.pdf} %{mashyp.pdf}
	\caption{Mapping Epistemological Concepts to (Cyber-Physical) Systems Engineering}
	\label{fig:mashyp}
\end{figure}

@@

\paragraph{Cybersecurity Weaknesses -- The RIDI-Hypothesis} 
It is important to note that all the results of the application of the
$\abftheory$-framework to channels (the analysis of the RCC relations
between output and input assertions of an agent)
%$\Rcc{\assertionRegion_o}{\assertionRegion_i}$) 
lead to the same results of the
analysis of a pair of an input and output port.
%(i.e.
%$\Rcc{\beliefRegion_1}{\assertionRegion_2}$ and
%$\Rcc{\assertionRegion_2}{\beliefRegion_3}$)
%As depicted in Figure~\ref{fig:atom}, 
So far we have
considered information generated by a port $P_I$ and then sent through a
channel $C$ to another (recipient) port $P_O$. In this scenario, where ports and
channels are atomic (otherwise raising infinite regress), we can only
consider the relations between ports and channel; considering both input-port
to channel and channel to output-port.  In fact, the weaknesses of a channel
are defined in terms of the weaknesses of ports.  
The same result can be obtained by analyzing the relation between the outputs
of a functional block and the inputs of another functional block, where
functional blocks are constituents of the functional architecture as described
afterwards.  In order to define a functional block without encountering an
infinitely recursive definition, we must reach the same conclusions as for the
channel. Therefore, describing the information as flowing over a channel or in
a functional block is purely syntactic sugar.

@@

%the insecurity of a system 
%($I$) can be expressed as the following equation.
%\begin{displaymath}
%	I=\prod_{a\in Ag} r_i^{\binom{n}{k}}-\delta_T
%\end{displaymath}
%where $r_i^{\binom{n}{k}}$ represents the number of possible configurations
%(permutations) of an agent (defined as a tuple of RCC relations as in
%Definition~\ref{def:opsystem}) in which at least one of the relation is not EQ.
%$r_i$ represents the RCC relations with arity $k$ between $n$ different regions
%in the $\abftheory$ theory. $\delta_T$ represents the number of configurations
%which are not satisfiable with respect to the logical theory defining the
%algebraic structure (topology) and constraints of the calculus (RCC). $Ag$ is
%the set of all the agents in the system under evaluation and the product
%represents all the unintended, insecure, configurations.  Given that there is
%only $1$ nominal configuration of the system where all the agents are defined
%only with EQ relations (and then the only secure configuration), which can be
%expressed as $sad$ it follows that security can be defined as
%\begin{displaymath}
%	I=\prod_{a\in Ag} r_i^{\binom{n}{k}}-\delta_T
%\end{displaymath}
%
%if $I=0$ the system can be considered secure, 
%otherwise $I$ represents the number of possible insecure configurations, where
%$I\in\mathbb{N}$.

@@

\paragraph{Case Studies}
We report the results of the evaluation of a water level reader (sensor ad-hoc example). 
\begin{figure}
	\centering
	\includegraphics[width=.5\columnwidth]{eng_cs1.png}
	\caption{Water level reader deployment diagram}
	\label{fig:eng_cs1}
\end{figure}
As in Figure~\ref{fig:eng_cs1}, we defined 2 agents: sensorInTank and sensorBoard,
which represent the physical reader that needs to be placed in a tank, and 
the board that interprets the readings and outputs them as signals.
The two components are connected by a wire.
\begin{figure}
	\centering
	\includegraphics[width=.9\columnwidth]{internal_cs1.png}
	\caption{Sensor Board Object Diagram}
	\label{fig:int_cs1}
\end{figure}
In Figure~\ref{fig:int_cs1}, we report the functional architecture that
receives the incoming communications from the sensor in the tank, and 
communicates them encrypted.
\begin{figure*}
	\centering
	\includegraphics[width=.7\textwidth]{results_excel.pdf}
	\caption{Partial View of the results in the spreadsheet file}
	\label{fig:results}
\end{figure*}
The results, summarized in Figure~\ref{fig:results}, reports 16777216
possible scenarios in which at least one component diverge from the specification.

%\section{Discussions}\label{sec:discussion}
%\paragraph{Abstraction Level}
%The level of abstraction is always a concern when it comes to the definition of
%a system model.  In this paper, we justified our assumptions on the atomicity
%of ports, channels, and functional blocks but nothing prevents a modeler to
%consider other atomic element of the system specification and design.  Those
%additional elements may detail the system and may change the syntactic sugar we
%defined; however, it won't change the underlying (semantic) reasoning unless
%another calculus is used (instead of RCC).  Furthermore, other Infosec
%properties may be considered, such as authenticity or non-repudiation. We,
%however, leave them as future development.
%
%\paragraph{$\abftheory$ Falsification} The theory we developed can be falsified
%by providing an experiment that shows a weaknesses not predicted by the
%$\abftheory$-framework, or by showing that the weaknesses predicted for a certain
%experiment are impossible to realize.  This theory, however, needs to be
%detailed to be applicable to a full engineering process (we only cover the
%first steps).  We propose the following complete process as a draft for
%potential next steps in this direction.
%\begin{enumerate}
%\item System Specification
%\item Architecture Design
%\item Security Risk Assessment and Security Requirement Extraction
%\item Design Refinement, where the semantics of the functional behavior and
%	communication protocols are defined 
%\item Security Requirement Extension, based on the design refinement
%\item Verification of Security Requirements (data flow, behaviors, \& protocol
%	logic)
%\item Software Prototype Implementation/Synthesis
%\item Automated Security Testing (abstract test-suite generation \&
%	concretization)
%\item Hardware Prototype Implementation
%\item Automated Security Testing
%\item Penetration Testing
%  \end{enumerate}
%
%\paragraph{Relation with DY}
%The DY attacker model has been extensively
%used for the identification of logical weaknesses in security protocols.
%Our theory should predict all possible weaknesses which, in turn,
%should enable all the attacks identified by the DY model. 
%A proof that the two theories/models correctly predicts
%(or lacks in predicting) weaknesses and related protocol flaws would
%connect risk assessment and protocol verification, the first
%two steps of the secure engineering of systems.

\section{Conclusion and Future Works}
%\fixnote{mv}{Se ho capito bene, detto a parole mie, quello che il vostro
%approccio permette di fare, dati requisiti e architettura del sistema, è di
%elencare in maniera sistematica tutti i "punti" in cui potrebbero sorgere dei
%problemi. Spiegherei un po' meglio, qui nelle conclusioni o nella sezione
%precedente, come questo approccio dovrebbe essere usato nella pratica... Voglio
%dire che non mi sembra necessariamente in opposizione ad altri approcci ma
%magari può integrarsi a questi.  Inoltre quando parlate dei passi successivi,
%ad esempio il considerare la generazione di test cases, se avete già qualche
%idea su come procedere, potreste discuterne brevemente.}
We proposed an hypothesis for a foundational theory on security, arguing that
cybersecurity-related issues are not linked to the maliciousness of an agent
but to the vagueness in the design processes.  We also provided a prototype
tool for the quantitative estimation of the cybersecurity risk, based on a UML
model of the system under assessment.
%The current state-of-the-art processes allow, given a specification,
%an engineer to design the system in such a way that cybersecurity issues arise due to the lack
%of proper cybersecurity risk assessment processes.
%Those designs, again, lack of cybersecurity verification processes based on a solid 
%cybersecurity foundational theory and then permit the generation of insecure implementation.
The cybersecurity verification and test-case generation will be the focus of our next steps.
%We conclude that the problem of security is a problem related to the 
%many possible design and, in turn, implementation given a specification.
%The problem is related to the epistemological search for truth,
%where the challenge is to relate information and belief to human knowledge.
%Scientifically, generalizing the human to an agent, the problem is
%to relate assertions and beliefs, to facts. From an engineering standpoint,
%by generalizing the concept of agents as architectural subsystems, the
%problem is to link channels (and ports) and functional architectures to 
%requirements.

%\section*{Acknowledgment}
%We thank Katia Santac\`a and Luca Vigan\`o for their help in the development of
%the initial ideas, and for the helpful discussions that allowed us to make this
%step forward. We also thank Marco Volpe for his valuable feedback.

@@

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

%\bibliographystyle{IEEEtran}
%\bibliography{../literature}

>>>>

Our literature review shows that most of the definitions relate insecurity to
dis-honesty (also called maliciousness or adversarial) of an agent (often
called adversary or attacker). In fact, weaknesses become vulnerability if an
attacker exploits them in an attack.  This, however, just shifts the problem of
defining what cybersecurity is as the problem of defining what a dishonest
agent can or cannot do in a system, where an agent is any virtual or physical
entity of the system or using the system (e.g. a device, a software, or a human
being), and dishonesty is not necessarily related to malicious motivation but
also to incompetence or lack of skills. As in the Dolev-Yao attacker model\footnote{The Dolev-Yao
attacker model can be considered as abstract model of an attacker that can be
used for the analysis of protocol specifications.} \autocite{Dolev1983security}, we may
correlate ``being dishonest'' to ``not following the intended behavior/rules''.
In the case of a generic system, a dishonest agent is, therefore, any agent
that doesn't follow the intended behavior (or functionality or logic) of the
system.  For example, a software can be considered an agent of the system and
whenever it has a bug, it can be exploited causing an incident. This
may lead to the conclusion that the dishonest behaviors of agents cannot be
defined in general (e.g. due to the heterogeneity of agents and systems) and
that a huge repository of dishonest behaviors should be kept as definition, such
as CAPEC \autocite{MITRE2020CAPEC}; or that dishonesty of an agent with respect
to a cybersecurity protocol should be defined as a number of predefined actions as
in
\autocite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}
(to name a few)\footnote{In formal protocol verifiers, in the DY model, the
attacker has a set of hard-coded inference rules that defines how the attacker
deduces new messages based on the observed message exchange.}.  As depicted in
Figure~\ref{fig:causality}, in order to define a theory on cybersecurity we may
say that the presence of vulnerabilities is a necessary condition in order to
cause an attack in the system.  Those vulnerabilities are, in turn, caused by
the presence of weaknesses in the system.  Weaknesses are errors in the design
or implementation of a system.  Therefore, a theory on cybersecurity should
first predict the errors in a system design.

@@

@@

@@

\subsection{From Multi-Agent to Cyber-Physical Systems}
\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth]{abftheory.pdf}
	\caption{Mapping Epistemological Concepts to (Cyber-Physical) Systems Engineering}
	\label{fig:mashyp}
\end{figure}

@@

\paragraph{Cybersecurity Weaknesses -- The RIDI-Hypothesis} 
It is important to note that all the results of the application of the
$\abftheory$-framework to channels (the analysis of the RCC relations
between output and input assertions of an agent)
lead to the same results of the
analysis of a pair of an input and output port.
So far we have
considered information generated by a port $P_I$ and then sent through a
channel $C$ to another (recipient) port $P_O$. In this scenario, where ports and
channels are atomic (otherwise raising infinite regress), we can only
consider the relations between ports and channel; considering both input-port
to channel and channel to output-port.  In fact, the weaknesses of a channel
are defined in terms of the weaknesses of ports.  
The same result can be obtained by analyzing the relation between the outputs
of a functional block and the inputs of another functional block, where
functional blocks are constituents of the functional architecture as described
afterwards.  In order to define a functional block without encountering an
infinitely recursive definition, we must reach the same conclusions as for the
channel. Therefore, describing the information as flowing over a channel or in
a functional block is purely syntactic sugar.

@@

@@

\paragraph{Case Studies}
We report the results of the evaluation of a water level reader (sensor ad-hoc example). 
\begin{figure}
	\centering
	\includegraphics[width=.5\columnwidth]{eng_cs1.png}
	\caption{Water level reader deployment diagram}
	\label{fig:eng_cs1}
\end{figure}
As in Figure~\ref{fig:eng_cs1}, we defined 2 agents: sensorInTank and sensorBoard,
which represent the physical reader that needs to be placed in a tank, and 
the board that interprets the readings and outputs them as signals.
The two components are connected by a wire.
\begin{figure}
	\centering
	\includegraphics[width=.9\columnwidth]{internal_cs1.png}
	\caption{Sensor Board Object Diagram}
	\label{fig:int_cs1}
\end{figure}
In Figure~\ref{fig:int_cs1}, we report the functional architecture that
receives the incoming communications from the sensor in the tank, and 
communicates them encrypted.
\begin{figure*}
	\centering
	\includegraphics[width=.7\textwidth]{results.pdf}
	\caption{Partial View of the results in the spreadsheet file}
	\label{fig:results}
\end{figure*}
The results, summarized in Figure~\ref{fig:results}, reports 16777216
possible scenarios in which at least one component diverge from the specification.

\section{Conclusion and Future Works}
We proposed an hypothesis for a foundational theory on security, arguing that
cybersecurity-related issues are not linked to the maliciousness of an agent
but to the vagueness in the design processes.  We also provided a prototype
tool for the quantitative estimation of the cybersecurity risk, based on a UML
model of the system under assessment.
The cybersecurity verification and test-case generation will be the focus of our next steps.

@@
